---
title: "Project"
output: html_document
---

# Chad and Tyler's initial work

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)

f_pvalue = function(p.value, symbol = "="){
  p.value <- round(p.value, digits = 3)
  if (p.value == 0) {
    return("p < .001")
  } else {
    return(paste0("p", symbol, round(p.value, digits = 3)))
  }
}
```

``` {r}
# Import raw data from source
wq.red <- read.csv("winequality-red.csv", sep=";")
wq.white <- read.csv("winequality-white.csv", sep=";")

category <- c(rep("red", nrow(wq.red)), rep("white", nrow(wq.white))) # create new column to show wine category

wq <- rbind(wq.red, wq.white) # concatenate red and white datasets
wq <- cbind(category, wq) # add category column
```

## Testing data summarization

```{r}
white.summary <- wq.white %>% summarise(mean.citric.acid = mean(citric.acid), sd.citric.acid = sd(citric.acid), median.citric.acid = median(citric.acid))
white.summary
```

The mean (standard deviation) citric acid content is `r round(white.summary$mean.citric.acid[1], digits = 2)` (`r round(white.summary$sd.citric.acid[1], digits = 2)`) <br>
The median citric acid content is `r round(white.summary$median.citric.acid[1], digits = 2)`.

## Test of variability pH

#### Hypotheses

* $H_0: \sigma_r = \sigma_w$
* $H_1: \sigma_r \ne \sigma_w$

#### Test statistic

```{r, warning=FALSE}
pHvar.f <- var.test(wq.red$pH, wq.white$pH, alternative = "t", conf.level = 0.95, ratio = 1)
```

* $F_0 = `r round(pHvar.f$statistic, digits = 2)`$.
* $`r f_pvalue(pHvar.f$p.value)`$.

#### Rejection region

Reject $H_0$ if $p < \alpha$, where $\alpha=0.05$. $`r f_pvalue(pHvar.f$p.value)` \nless 0.05.$

#### Conclusion and interpretation

Fail to reject $H_0$. There is insufficient evidence to suggest that there is a difference in the variability of pH between red and white wines.

# Jovan's initial work

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
#library(car)
#library(pander)
#library(agricolae)
#library(DescTools)
library(stats)
library(BSDA)
#library(pgirmess)
#library(multcomp)

### p.value.string v2
# Update v2: added the formatting that turns of scientific notation
# fixes the case when p = 0.0001 (instead of p=1e-4)
# This function called p.value.string creates a string
# to be used when reporting the p-value. It includes the p on the LHS.
# You will likely place it in $$ to get the LaTeX math formatting.
# This should be placed in an r code chunk in the beginning of your R markdown
# This code chunk by itself shouldn't produce any result.


p.value.string = function(p.value){
  p.value <- round(p.value, digits=4)
  if (p.value == 0) {
    return("p < 0.0001")
  } else {
    return(paste0("p = ", format(p.value, scientific = F)))
  }
}

knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
f_pvalue = function(p.value, symbol = "="){
  p.value <- round(p.value, digits = 3)
  if (p.value == 0) {
    return("p < .001")
  } else {
    return(paste0("p", symbol, round(p.value, digits = 3)))
  }
}

```



```{r}
redwine <- read.csv("winequality-red.csv", sep=";")
whitewine <- read.csv("winequality-white.csv", sep=";")
wine <- rbind(redwine, whitewine)
type <- as.factor(c(rep("Red Wine", nrow(redwine)), rep("White Wine", nrow(whitewine))))
wine <- cbind(wine, type)
```


## 1. One Way Anova Test on the contain of Sulfure Dioxide

```{r}
wine.Summary <-  wine %>% group_by(type) %>% summarise(mean_alcohol = mean(total.sulfur.dioxide), stdev_alcohol = sd(total.sulfur.dioxide))
wine.Summary
```

```{r}
wine_anova_results <- aov(total.sulfur.dioxide ~ type, data = wine)
wine_anova_table <- summary(wine_anova_results)
wine_anova_table
#pander(wine_anova_table, style='rmarkdown') 
```

#### Hypothesis test

**Hypotheses**

$H_0: \ \mu_1 = \mu_2$ <br> 
$H_1: \ \mu_1 \ne \mu_2$

**Test Statistic**

$F_0 = `r p.value.string(wine_anova_table[[1]][["F value"]][1])`$.

***p*-value**

$`r p.value.string(wine_anova_table[[1]][["Pr(>F)"]][1])`$.

**Rejection Region**

Reject if $p < \alpha$, where $\alpha=0.05$.

**Conclusion and Interpretation**

Reject $H_0$. There is sufficient evidence to suggest that the total sulfure dioxide contain in the two wine are diferents.


## 2. Kruskal-Wallis Test on the contain of Alcohol

```{r}
wine.Summary2 <-  wine %>% group_by(type) %>% summarise(alcohol_median = median(alcohol), alcohol_IQR = IQR(alcohol))
wine.Summary2
```

```{r}
wine_KW_results <- kruskal.test(alcohol ~ type, data = wine)
wine_KW_results
```

**Hypotheses**

$H_0: \ \mbox{M}_1 = \mbox{M}_2$ <br> 
$H_1: \ \mbox{M}_1 \ne \mbox{M}_2$ 

**Test Statistic**

$H = `r round(wine_KW_results$statistic,digits=2)`$.

***p*-value**

$`r p.value.string(wine_KW_results$p.value)`$.

**Rejection Region**

Reject if $p < \alpha$, where $\alpha=0.05$.

**Conclusion and Interpretation**

Fail to reject $H_0$. There is not sufficient evidence to suggest that there is a diferent in the alcohol contain of the two type of wines.


## 3. Test of variability pH

```{r}
pHvar <- var.test(redwine$pH, whitewine$pH, alternative = "t", conf.level = 0.95, ratio = 1)
pHvar 
```

**Hypotheses**

* $H_0: \sigma_r = \sigma_w$
* $H_1: \sigma_r \ne \sigma_w$

**Test Statistic**

* $F_0 = `r round(pHvar$statistic, digits = 2)`$.
* $`r f_pvalue(pHvar$p.value)`$.

**Rejection Region**

Reject $H_0$ if $p < \alpha$, where $\alpha=0.05$. $`r f_pvalue(pHvar$p.value)` \nless 0.05.$

**Conclusion and Interpretation**

Fail to reject $H_0$. There is insufficient evidence to suggest that there is a difference in the variability of pH between red and white wines.


## 4. Test of contain of Chlorides

Formally test to determine if the mean of the chlorides contain on red wine is greater than that of white wine. Use the appropriate version of the t-test with the critical at the α = 0.05 level.

```{r}
wine.Summary <-  wine %>% group_by(type) %>% summarise(mean_alcohol = mean(chlorides), stdev_alcohol = sd(chlorides))
wine.Summary
```

```{r}
chlorides_t <- t.test(redwine$chlorides, whitewine$chlorides, paired = FALSE, alternative = "greater")
chlorides_t
chlorides_p <- p.value.string(chlorides_t$p.value)
```

**Hypotheses**

$H_0: \ \mu_R \le \mu_W$ <br> 
$H_1: \ \mu_R > \mu_w$

**Test Statistic**

$t_0 = `r round(chlorides_t$statistic, digits=3)`$.

***p*-value**

$`r chlorides_p`$.

**Rejection Region**

Reject if $p < \alpha$, where $\alpha=0.01$.

**Conclusion and Interpretation**

Reject $H_0$. There is sufficient evidence to suggest that the contain chlorides in the red wine are greather than the white wine.


## 5. Test of variability of residual sugar 

```{r}
rsvar <- var.test(redwine$residual.sugar, whitewine$residual.sugar, alternative = "t", conf.level = 0.95, ratio = 1)
rsvar
```

**Hypotheses**

* $H_0: \sigma_r = \sigma_w$
* $H_1: \sigma_r \ne \sigma_w$

**Test Statistic**

* $F_0 = `r round(rsvar$statistic, digits = 2)`$.
* $`r f_pvalue(rsvar$p.value)`$.

**Rejection Region**

Reject $H_0$ if $p < \alpha$, where $\alpha=0.05$. $`r f_pvalue(rsvar$p.value)` < 0.05.$

**Conclusion and Interpretation**

Reject $H_0$. There is sufficient evidence to suggest that there is a difference in the variability of residual sugar contain between red and white wines.


## 6. Test of contain of Citrid Acid

Formally test to determine if the mean of the citrid acid contain on red wine is less than that of white wine. Use the appropriate version of the t-test with the critical at the α = 0.05 level.

```{r}
wine.Summary3 <-  wine %>% group_by(type) %>% summarise(mean_alcohol = mean(citric.acid), stdev_alcohol = sd(citric.acid))
wine.Summary3
```

```{r}
citricacid_t <- t.test(redwine$citric.acid, whitewine$citric.acid, paired = FALSE, alternative = "less")
citricacid_t
citricacid_p <- p.value.string(chlorides_t$p.value)
```

**Hypotheses**

$H_0: \ \mu_R \ge \mu_W$ <br> 
$H_1: \ \mu_R < \mu_w$

**Test Statistic**

$t_0 = `r round(citricacid_t$statistic, digits=3)`$.

***p*-value**

$`r citricacid_p`$.

**Rejection Region**

Reject if $p < \alpha$, where $\alpha=0.01$.

**Conclusion and Interpretation**

Reject $H_0$. There is sufficient evidence to suggest that the contain of citric acid in the red wine is less than the white wine.


## 7. Proportion Test on Density??? 

```{r}
wine_density1 <- z.test(wine$density, alternative = "greater", mu = 1, sigma.x = sd(wine$density))
wine_density1
```

**Hypotheses**

$H_0: \ \mu \le 1$ <br>
$H_1: \ \mu > 1$

**Test Statistic**

$z = `r wine_density1$statistic`$.

***p*-value**

$p = `r wine_density1$p.value`$.

**Rejection Region**

Reject if $p < \alpha$, where $\alpha=0.05$.

**Conclusion and Interpretation**

Fail to reject $H_0$. There is not sufficient evidence to suggest that the mean density wine is greater than 1.

```{r}
wine_density2 <- z.test(wine$density, alternative = "less", mu = 0.98, sigma.x = sd(wine$density))
wine_density2
```

**Hypotheses**

$H_0: \ \mu \ge 0.98$ <br>
$H_1: \ \mu > 0.98$

**Test Statistic**

$z = `r wine_density2$statistic`$.

***p*-value**

$p = `r wine_density2$p.value`$.

**Rejection Region**

Reject if $p < \alpha$, where $\alpha=0.05$.

**Conclusion and Interpretation**

Fail to reject $H_0$. There is not sufficient evidence to suggest that the mean density wine is less than 0.98.

```{r}
wine_density3 <- z.test(wine$density, alternative = "less", mu = 0.99, sigma.x = sd(wine$density))
wine_density3
```

**Hypotheses**

$H_0: \ \mu \ge 0.99$ <br>
$H_1: \ \mu > 0.99$

**Test Statistic**

$z = `r wine_density3$statistic`$.

***p*-value**

$p = `r wine_density3$p.value`$.

**Rejection Region**

Reject if $p < \alpha$, where $\alpha=0.05$.

**Conclusion and Interpretation**

Fail to reject $H_0$. There is not sufficient evidence to suggest that the mean density wine is less than 0.99.


## 8. Regression model

```{r}
one_model <- lm(quality ~ fixed.acidity + volatile.acidity + citric.acid + residual.sugar + chlorides + total.sulfur.dioxide + density + pH + sulphates + alcohol, data=wine)
one_coef <- coefficients(one_model)
one_anova <- anova(one_model)
one_summary <- summary(one_model)
one_t <- as_tibble(one_summary[[4]])
one_ci <- as_tibble(confint(one_model, level=0.95))
```

The resulting regression model is 
\[ \hat{y} = `r one_coef[[1]]` + `r one_coef[[2]]`x_1 + `r one_coef[[3]]`x_2 + `r one_coef[[4]]`x_3 + `r one_coef[[5]]`x_4 + `r one_coef[[6]]`x_5 + `r one_coef[[7]]`x_6 + `r one_coef[[8]]`x_7 + `r one_coef[[9]]`x_8 + `r one_coef[[10]]`x_9\]

### ANOVA table and testing for significance of regression line

The corresponding ANOVA table is as follows

```{r}
one_anova
```

**Hypotheses**

$H_0: \ \beta_1 = 0$ <br>
$H_1: \ \beta_1 \ne 0$

**Test Statistic**

$F_0 = `r round(one_anova$"F value"[1], digits=2)`$.

***p*-value**

$`r p.value.string(one_anova$"Pr(>F)"[3])`$.

**Rejection Region**

Reject if $p < \alpha$, where $\alpha=0.05$.

**Conclusion and Interpretation**

Comparing every p value of the previous table with the alpha, can be check that with at lest one of then is greather than alpha, by this reason fail to reject $H_0$. There is not sufficient evidence to suggest that the regression line is significant.

### Testing for significance of individual predictor

```{r}
one_t
```

**Hypotheses**

$H_0: \ \beta_1 = 0$ <br>
$H_1: \ \beta_1 \ne 0$

**Test Statistic**

$t_0 = `r round(one_t$"t value"[2], digits=2)`$.

***p*-value**

$`r p.value.string(one_t$"Pr(>|t|)"[4])`$.

**Rejection Region**

Reject if $p < \alpha$, where $\alpha=0.05$.

**Conclusion and Interpretation**

Comparing every p value of the previous table with the alpha, can be check that with at lest one of then is greather than alpha, by this reason fail to reject $H_0$. There is not sufficient evidence to suggest that the regression line is significant.

