---
title: 'Project: Wine Quality'
author:
- Jovan Burgos
- Adrienne Eddy
- Chad Michael Eyer
- Valorie Myer
- Tyler Watson
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  html_document:
    includes:
      after_body: doc_prefix.html
    theme: cerulean
  pdf_document: default
nocite: |
  @Cortez2009aa
csl: apa.csl
bibliography: project.bib
---

<!--Detailed List of Contributions
***On the final exam Dr. Seals will ask you to name everyone in your group along with all of their contributions. Please add your contributions you've done for this project and then get a rough idea of what everyone did before the final.

Jovan Burgos:
Created a code for an assortment of tests. Created idea of having wine quality for data set.Gave results for test three, four, and five

Adrienne Eddy:
Completed the methods section. Worked on introduction. Worked on hypothesis 5, created idea for hypothesis six. worked on code for tests for hypothesis two, six. Gave results for test three, four and six. Worked on cleaning up project.

Chad Michael Eyer:
Created and trained on GitHub. Hosted virtual meeting. Created hypotheses. Completed test for hypotheses one, two, and three. Found concrete data set. General clean-up. Biblography work. Created set-up block.

Valorie Myer:
Created group chat for the group. Completed introduction. Hosted virtual meeting. Wrote conclusions section. Rewrote results as a paragraph. Worked on cleaning up project.

Tyler Watson:
Formulated hypotheses. Created detailed list of contributions. Completed test for hypotheses one and two. Put methods section into past tense. Completed test for hypothesis five.

-->

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(stats)
library(BSDA)
library(pander)
library(alr3)
library(knitr)

p.value.string = function(p.value, symbol = "="){
  p.value <- round(p.value, digits = 4)
  if (p.value == 0) {
    return("p < .001")
  } else {
    return(paste0("p", symbol, round(p.value, digits = 4)))
  }
}

#### Import raw data from source
wine_red <- read.csv("winequality-red.csv", sep=";")
wine_white <- read.csv("winequality-white.csv", sep=";")

type <- c(rep("Red Wine", nrow(wine_red)), rep("White Wine", nrow(wine_white))) # create new column to show wine category

wine <- rbind(wine_red, wine_white) # concatenate red and white datasets
wine <- cbind(type, wine) # add category column
```

## Introduction

Wine testing is important to both sellers and consumers. By testing wine samples, vineyards are able to ensure that their wines are safe for consumers, and that the taste of each type of wine stays consistent. It is even used to help set pricing scales. This project will be comparing different physiochemical (chemical properties) and sensory elements of wine in order to see what elements impact quality.

The chemical composition of wine can impact both the quality of the wine and our bodies. This study will be looking at 11 different physiochemical properties: fixed acidity, volatile acidity, citric acid, residual sugar, chlorides, free sulfur dioxide, total sulfur dioxide, density, pH, sulphates, and alcohol content. We will specifically look to see if there is an interaction between volatile acidity and citric acidity. Volatile acidity is how wine spoilage is measured [@siteC]. Typically, wine makers attempt to keep volatile acidity to a minimum to improve the quality of wine [@siteA]. Citric acid keeps wine tasting fresh, while imparting a slight citric taste to the wine [@siteB].

## Hypotheses

This project answers the following research questions:

- Are variances of quality of red wines equal to variances of quality of white wines?
- Are mean quality of red wines equal to mean quality of white wines?
- Do physiochemical attributes significantly affect quality?
- If so which physiochemical attributes significantly affect quality?
- Is there an interaction between citric acid and volatile acidity for red wines?
- What physiochemical attribute differences might lead to better quality?

## Methods

All tests are run at the $95\%$ confidence level unless otherwise noted.

Below, you will find the tests for each hypothesis and a brief explanation and justification for each test.

*Variances of quality*

We used the $F$ test because this test allows us to compare the variances of each data set to see if they are equal. We used this to guide us in determining which tests to use along the way.

*Mean quality*

Provided that the data met the assumptions, we used $t$-tests because the two sets of data are independent and this test allows us to compare means.

*Do any physiochemical attributes affect quality?*

We used multivariable regression and a multivariable ANOVA because this allowed us to create a model and analyze this model that will showed if there is any relationship between any of the attributes and the quality of the wine.  We created pairwise scatterplots which give us a visual idea of what the data looks like.  

*Which physiochemical attributes affect quality?*

We used the results from the multivariable ANOVA from the previous hypothesis to determine which attributes were significant predictors of quality.

*Is there an interaction between citric acid and volatile acidity?*

We created a regression model to determine if there is an interaction between citric acid and volatile acidity. 

*Physiochemical attributes potentially improving quality*

We ran t-tests after testing assumptions to compare physiochemical attributes between red and white wine to determine the kind of difference.


## Results

This project will compares different physiochemical and sensory elements of wine. The wine in this data set are Vinho Verde wines, which comes from Minho region of Portugal. The data was collected between May 2004 and February 2007 and includes 1599 red wine samples and 4898 white wine samples. This study looked at 11 different physiochemical variables: fixed acidity, volatile acidity, citric acid, residual sugar, chlorides, free sulfur dioxide, total sulfur dioxide, density, pH, sulphates, and alcohol content. This data also contains a sensory score variable called “quality.” Each wine sample was tasted by at least 3 judges, each of whom scored the wine from 0 (very bad) to 10 (excellent). The quality variable was the median of the scores [@Cortez2009aa]. 

``` {r , include=FALSE}
###Variance (Hypothesis 1)

wine_quality_var.f <- var.test(wine_red$quality, wine_white$quality, alternative = "t", conf.level = 0.95, ratio = 1)


wine_anova <- aov(quality ~ type, data = wine)
wine_anova_table <- summary(wine_anova)
wine_anova_table

plot_anova_assumptions <- function(aov.results){
  par(mfrow=c(2,2))
  plot(aov.results, which=1)
  plot(aov.results, which=2)
  aov_residuals <- residuals(aov.results)
  plot(density(aov_residuals))
  hist(aov_residuals)
}

plot_anova_assumptions(wine_anova)


winequal <- t.test(wine_red$quality, wine_white$quality, var.equal = FALSE)
winequal
winequal.t <- winequal$statistic[1]
winequal.t
winequal.p <- winequal$p.value[1]
winequal.p

###Mean Quality (Hypothesis 2)
winequal1 <- t.test(wine_red$quality, wine_white$quality, var.equal = FALSE, alternative = "greater")
winequal1

winequal2 <- t.test(wine_red$quality, wine_white$quality, var.equal = FALSE, alternative = "less")
winequal2

### Physiochemical attributes and quality (Hypothesis 3 and 4)
```
``` {r, include= FALSE}
h3_model <- lm(quality ~ fixed.acidity + volatile.acidity + citric.acid + residual.sugar + chlorides + free.sulfur.dioxide + total.sulfur.dioxide + density + pH + sulphates + alcohol, data=wine)
h3_model.coef <- coefficients(h3_model)
h3_anova <- Anova(h3_model, type = 3, digits = 3)
h3_summary <- summary(h3_model)

## Formatting ANOVA table...
h3_anova <- h3_anova[-c(1),] # Remove intercept row
h3_anova$MS <- h3_anova[,1]/h3_anova[,2] # Add column for mean sum of squares
h3_total_row <- apply(h3_anova, 2, sum) # Create "Total" row
h3_total_row.MS <- NA # Set mean sum of squares on "Total" row to NA
h3_anova <- rbind(h3_anova,h3_total_row) # Add "Total" row to bottom of ANOVA table
h3_anova <- cbind(h3_anova[c(1,2,5,3,4)]) # Reorder columns of ANOVA table
h3_anova$`Pr(>F)` <- ifelse(h3_anova$`Pr(>F)` < .0001, "< 0.0001", format(round(h3_anova$`Pr(>F)`, digits = 4), nsmall = 4)) # Change display of p values so they aren't rounded to zero
colnames(h3_anova) <- c("SS", "df", "MS", "$F$", "$p$") # Set readable column headings for ANOVA table
rownames(h3_anova) <- c("Fixed acidity", "Volatile acidity", "Citric acid", "Residual sugar", "Chlorides", "Free sulfur dioxide", "Total sulfur dioxide", "Density", "pH", "Sulphates", "Alcohol", "Error", "Total") # Set readable row headings for ANOVA table
options(knitr.kable.NA = '') # this will hide missing values in the kable table


kable(h3_anova, digits = 3, format = "pandoc", caption = "ANOVA table", align=c('l', rep('r', 5))) # Output ANOVA table


### Which physiochemical attributes (Hypothesis 4 graph)


pairs(~sulphates+total.sulfur.dioxide+chlorides+citric.acid+pH+volatile.acidity+residual.sugar+fixed.acidity+density+alcohol, data=wine)

### Citric acid and volatile acidity (Hypothesis 5)

```{r, include= FALSE}
wine_red_results <- aov(quality ~ citric.acid*volatile.acidity, data = wine_red )

wine_red_table <- summary(wine_red_results)

pander(wine_red_table, style='rmarkdown') 

### Which PC impact quality (Hypothesis 6)
compare1 <- t.test(wine_white$fixed.acidity, wine_red$fixed.acidity, var.equal = FALSE, alternative = "less")
```

```{r, include= FALSE}
compare2 <- t.test(wine_white$volatile.acidity, wine_red$volatile.acidity, var.equal = FALSE, alternative = "less")
```

```{r, include= FALSE}
compare3 <- t.test(wine_white$residual.sugar, wine_red$residual.sugar, var.equal = FALSE, alternative = "greater")
```

```{r, include= FALSE}
compare4 <- t.test(wine_white$free.sulfur.dioxide, wine_red$free.sulfur.dioxide, var.equal = FALSE, alternative = "greater")
```

```{r, include= FALSE}
compare5 <- t.test(wine_white$total.sulfur.dioxide, wine_red$total.sulfur.dioxide, var.equal = FALSE, alternative = "greater")
```

```{r, include= FALSE}
compare6 <- t.test(wine_white$density, wine_red$density, var.equal = FALSE, alternative = "less")
```

```{r, include= FALSE}
compare7 <- t.test(wine_white$pH, wine_red$pH, var.equal = TRUE, alternative = "less")
```

```{r, include= FALSE}
compare8 <- t.test(wine_white$sulphates, wine_red$sulphates, var.equal = FALSE, alternative = "less")
```

```{r, include= FALSE}
compare9 <- t.test(wine_white$alcohol, wine_red$alcohol, var.equal = FALSE, alternative = "greater")
```

The first research question asked if the variance of quality was the same for both red and white wines. There is sufficient evidence to suggest that the variability of the quality scores of red and white wines is significantly different, with ($`r p.value.string(wine_quality_var.f$p.value)`$) at the $\alpha = .05$ level. Because there is a difference in variability, it impacted which tests were run in the following hypothesis tests. 


The next research question was to determine if the mean quality of the two types of wines were different. There is sufficient evidence to suggest that the mean quality is different, with ($`r p.value.string(wine_anova_table[[1]][["Pr(>F)"]][1])`$) at the $\alpha = .05$ level. 

The natural follow-up question is, "Which wine has a greater mean quality" From several t-tests, it was determined that there is sufficient evidence to suggest that the mean quality of red wine is less than the mean quality of white wine at the $\alpha=.05$ level ($`r p.value.string(winequal2$p.value[1])`$). 

The next research question was to see if physiochemical properties affect the quality of wine. The resulting multivariable regression line is The resulting regression model is 

$\hat{y} = `r round(h3_model.coef[[1]], digits = 2)` + `r round(h3_model.coef[[2]], digits = 2)`$ fixed acidity  $f+`r round(h3_model.coef[[3]], digits = 2)`$ volatile acidity $+ `r round(h3_model.coef[[4]], digits = 2)`$ citric acid $+ `r round(h3_model.coef[[5]], digits = 2)`$ residual sugar $+ `r round(h3_model.coef[[6]], digits = 2)`$ chlorides $+ `r round(h3_model.coef[[7]], digits = 2)`$ free sulfur dioxide $+ `r round(h3_model.coef[[8]], digits = 2)`$ tota sulfur dioxide $+ `r round(h3_model.coef[[9]], digits = 2)`$ density $+ `r round(h3_model.coef[[10]], digits = 2)`$ pH $+ `r round(h3_model.coef[[11]], digits = 2)`$ sulphates $+ `r round(h3_model.coef[[12]], digits = 2)`$ alcohol


The multivariable ANOVA test (results shown below, along with pairwise comparisons of the different physiochemical attributes) shows that there is sufficient evidence to suggest that wine quality scores are significantly affected by fixed acidity, volatile acidity, residual sugar, free sulfur dioxide, total sulfur dioxide, density, pH, sulphates, and alcohol. Quality scores are not dependent on citric acid or chlorides.

```{r}
kable(h3_anova, digits = 3, format = "pandoc", caption = "ANOVA table",
      align=c('l', rep('r', 5))) # Output ANOVA table
pairs(~ sulphates + total.sulfur.dioxide + 
        chlorides + citric.acid + pH + 
        volatile.acidity + residual.sugar + 
        fixed.acidity + density +
        alcohol, data=wine)
```


```{r}
wine %>% 
  ggplot(aes(x=wine$fixed.acidity, y=wine$quality
            )) +
    geom_point() +
  labs(title="quality compared to fixed acidity")+
  xlab("fixed acidity")+
  ylab("quality")+
  theme_bw() 
```
```{r}
wine %>% 
  ggplot(aes(x=wine$volatile.acidity, y=wine$quality
            )) +
    geom_point() +
  labs(title="quality compared to volatile acidity")+
  xlab("volatile acidity")+
  ylab("quality")+
  theme_bw() 
```

```{r}
wine %>% 
  ggplot(aes(x=wine$citric.acid, y=wine$quality
            )) +
    geom_point() +
  labs(title="quality compared to citric acid")+
  xlab("citric acid")+
  ylab("quality")+
  theme_bw() 
```

```{r}
wine %>% 
  ggplot(aes(x=wine$residual.sugar, y=wine$quality
            )) +
    geom_point() +
  labs(title="quality compared to residual sugar")+
  xlab("residual sugar")+
  ylab("quality")+
  theme_bw() 
```

```{r}
wine %>% 
  ggplot(aes(x=wine$chlorides, y=wine$quality
            )) +
    geom_point() +
  labs(title="quality compared to chlorides")+
  xlab("chlorides")+
  ylab("quality")+
  theme_bw() 
```

```{r}
wine %>% 
  ggplot(aes(x=wine$free.sulfur.dioxide, y=wine$quality
            )) +
    geom_point() +
  labs(title="quality compared to free sulfur dioxide")+
  xlab("free sulfur dioxide")+
  ylab("quality")+
  theme_bw() 
```

```{r}
wine %>% 
  ggplot(aes(x=total.sulfur.dioxide, y=wine$quality
            )) +
    geom_point() +
  labs(title="quality compared to total sulfur dioxide")+
  xlab("total sulfur dioxide")+
  ylab("quality")+
  theme_bw() 
```

```{r}
wine %>% 
  ggplot(aes(x=wine$density, y=wine$quality
            )) +
    geom_point() +
  labs(title="quality compared to density")+
  xlab("density")+
  ylab("quality")+
  theme_bw() 
```

```{r}
wine %>% 
  ggplot(aes(x=wine$pH, y=wine$quality
            )) +
    geom_point() +
  labs(title="quality compared to pH")+
  xlab("pH")+
  ylab("quality")+
  theme_bw() 
```

```{r}
wine %>% 
  ggplot(aes(x=wine$sulphates, y=wine$quality
            )) +
    geom_point() +
  labs(title="quality compared to sulphates")+
  xlab("sulphates")+
  ylab("quality")+
  theme_bw() 
```

```{r}
wine %>% 
  ggplot(aes(x=wine$alcohol, y=wine$quality
            )) +
    geom_point() +
  labs(title="quality compared to alcohol")+
  xlab("alcohol")+
  ylab("quality")+
  theme_bw() 
```
Winemakers use citric acid to sterilize equipment between batches, and it is found in fruitier wines.  However, citric acid can contribute to volatile acidity which leads to wine spoilage. Even though citric acid is not a significant predictor of quality, we want to see if there is an interaction between citric acid and volatile acidity.There is sufficient evidence to suggest that there is an interaction between citric acid and volatile acidity because $`r p.value.string(wine_red_table[[1]][["Pr(>F)"]][3])`$.

Knowing that all the physiochemical properties except for citric acid or chlorides these are all significant predictors of quality and that white wine has the higher quality, the last research question was to see how the physiochemical composition differs between white and red wines. There is sufficient evidence to suggest that white wine has less fixed acidity ($`r p.value.string(compare1$p.value[1])`$), less volatile acidity (spoilage) ($`r p.value.string(compare2$p.value[1])`$), more residual sugar ($`r p.value.string(compare3$p.value[1])`$), greater free sulfur dioxide ($`r p.value.string(compare4$p.value[1])`$), greater total sulfur dioxide ($`r p.value.string(compare5$p.value[1])`$), less dense ($`r p.value.string(compare6$p.value[1])`$), lower pH ($`r p.value.string(compare7$p.value[1])`$), fewer sulphates ($`r p.value.string(compare8$p.value[1])`$), and more alcohol ($`r p.value.string(compare9$p.value[1])`$).

## Conclusion

This study of Portuguese wines shows that winemaking is both an art and a science. As vineyards strive to develop better tasting wines, they need to consider the physiochemical make-up of their product. The study found that the physiochemical properties of wine play an important part in the quality. Out of the eleven physiochemical properties studied, all but two, citric acid and chlorides, significantly predicted the quality of the wine. The data from this study showed that the white wine quality was better than red wine quality. When we looked at how the red and white wines differed in physiochemical properties, we found that white wine has less fixed acidity, less volatile acidity (spoilage), more residual sugar, greater free sulfur dioxide, greater total sulfur dioxide, less dense, lower pH, fewer sulphates, and more alcohol. Another notable find is that these physiochemical properties don’t just impact the quality of the wines--they also impact each other, as seen by the interaction between citric acid and volatile acidity.  These are important concepts to keep in mind for large vineyards all the way to those who want to try making their own wine at home.
 
One of the limitations of this study was that all the wine was from the same region of Portugal. Another limitation was that the quality variable was based on a score of small groups of judges--only 3 judges in some cases. Something important to keep in mind when considering the quality score is that the score is based on the median, rather than the average, of the judges' scores. Future studies could be done to include wines from different parts of the globe and to include a larger panel of judges.

## Session Info

```{r sessionInfo, include=TRUE, echo=TRUE, results='markup'}
sessionInfo()
```

## References
