---
title: "Project"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)

f_pvalue = function(p.value, symbol = "="){
  p.value <- round(p.value, digits = 3)
  if (p.value == 0) {
    return("p < .001")
  } else {
    return(paste0("p", symbol, round(p.value, digits = 3)))
  }
}
```

``` {r}
# Import raw data from source
wq.red <- read.csv("winequality-red.csv", sep=";")
wq.white <- read.csv("winequality-white.csv", sep=";")
```

## Testing data summarization

```{r}
white.summary <- wq.white %>% summarise(mean.citric.acid = mean(citric.acid), sd.citric.acid = sd(citric.acid), median.citric.acid = median(citric.acid))
white.summary
```

The mean (standard deviation) citric acid content is `r round(white.summary$mean.citric.acid[1], digits = 2)` (`r round(white.summary$sd.citric.acid[1], digits = 2)`) <br>
The median citric acid content is `r round(white.summary$median.citric.acid[1], digits = 2)`.

## Test of variability pH

#### Hypotheses

* $H_0: \sigma_r = \sigma_w$
* $H_1: \sigma_r \ne \sigma_w$

#### Test statistic

```{r, warning=FALSE}
pHvar.f <- var.test(wq.red$pH, wq.white$pH, alternative = "t", conf.level = 0.95, ratio = 1)
```

* $F_0 = `r round(pHvar.f$statistic, digits = 2)`$.
* $`r f_pvalue(pHvar.f$p.value)`$.

#### Rejection region

Reject $H_0$ if $p < \alpha$, where $\alpha=0.05$. $`r f_pvalue(pHvar.f$p.value)` \nless 0.05.$

#### Conclusion and interpretation

Fail to reject $H_0$. There is insufficient evidence to suggest that there is a difference in the variability of pH between red and white wines.
