---
title: "Project: Wine Quality"
author: 
- "Jovan Burgos"
- "Adrinne Eddy"
- "Chad Michael Eyer"
- "Valorie Myer"
- "Tyler Watson"
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  html_document:
    includes:
      after_body: doc_prefix.html
    theme: cerulean
nocite: |
  @Cortez2009aa
csl: apa.csl
bibliography: project.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(stats)
library(BSDA)

p.value.string = function(p.value, symbol = "="){
  p.value <- round(p.value, digits = 3)
  if (p.value == 0) {
    return("p < .001")
  } else {
    return(paste0("p", symbol, round(p.value, digits = 3)))
  }
}

#### Import raw data from source
wine_red <- read.csv("winequality-red.csv", sep=";")
wine_white <- read.csv("winequality-white.csv", sep=";")

type <- c(rep("Red Wine", nrow(wine_red)), rep("White Wine", nrow(wine_white))) # create new column to show wine category

wine <- rbind(wine_red, wine_white) # concatenate red and white datasets
wine <- cbind(type, wine) # add category column
```

## Introduction

Briefly give meaning to your project – what is the background and why is this project important?  (20 points)

## Hypotheses

Clearly state the hypotheses being tested.  (20 points)

## Methods

State and explain the methodology used to test the hypotheses.  (20 points)

## Results

Describe the data (1st paragraph) as well as analysis results (order of results should correspond to hypotheses listed).  (20 points)

## Conclusion

Restate the results in terms of the larger picture; state limitations and opportunities for future research (20 points)

## Reference

<!-- Work

### Chad and Tyler's initial work

#### Testing data summarization

```{r}
white.summary <- wine_white %>% summarise(mean.citric.acid = mean(citric.acid), sd.citric.acid = sd(citric.acid), median.citric.acid = median(citric.acid))
white.summary
```

The mean (standard deviation) citric acid content is `r round(white.summary$mean.citric.acid[1], digits = 2)` (`r round(white.summary$sd.citric.acid[1], digits = 2)`) <br>
The median citric acid content is `r round(white.summary$median.citric.acid[1], digits = 2)`.

#### Test of variability pH

##### Hypotheses

* $H_0: \sigma_r = \sigma_w$
* $H_1: \sigma_r \ne \sigma_w$

##### Test statistic

```{r, warning=FALSE}
pHvar.f <- var.test(wine_red$pH, wine_white$pH, alternative = "t", conf.level = 0.95, ratio = 1)
```

* $F_0 = `r round(pHvar.f$statistic, digits = 2)`$.
* $`r p.value.string(pHvar.f$p.value)`$.

##### Rejection region

Reject $H_0$ if $p < \alpha$, where $\alpha=0.05$. $`r p.value.string(pHvar.f$p.value)` \nless 0.05.$

##### Conclusion and interpretation

Fail to reject $H_0$. There is insufficient evidence to suggest that there is a difference in the variability of pH between red and white wines.

### Jovan's initial work

```{r}
# wine_red <- read.csv("winequality-red.csv", sep=";")
# wine_white <- read.csv("winequality-white.csv", sep=";")
# wine <- rbind(wine_red, wine_white)
# type <- as.factor(c(rep("Red Wine", nrow(wine_red)), rep("White Wine", nrow(wine_white))))
# wine <- cbind(wine, type)
```

#### 1. One Way Anova Test on the contain of Sulfure Dioxide

```{r}
wine.Summary <-  wine %>% group_by(type) %>% summarise(mean_alcohol = mean(total.sulfur.dioxide), stdev_alcohol = sd(total.sulfur.dioxide))
wine.Summary
```

```{r}
wine_anova_results <- aov(total.sulfur.dioxide ~ type, data = wine)
wine_anova_table <- summary(wine_anova_results)
wine_anova_table
#pander(wine_anova_table, style='rmarkdown') 
```

##### Hypothesis test

**Hypotheses**

$H_0: \ \mu_1 = \mu_2$ <br> 
$H_1: \ \mu_1 \ne \mu_2$

**Test Statistic**

$F_0 = `r p.value.string(wine_anova_table[[1]][["F value"]][1])`$.

***p*-value**

$`r p.value.string(wine_anova_table[[1]][["Pr(>F)"]][1])`$.

**Rejection Region**

Reject if $p < \alpha$, where $\alpha=0.05$.

**Conclusion and Interpretation**

Reject $H_0$. There is sufficient evidence to suggest that the total sulfure dioxide contain in the two wine are diferent.

### Adrienne's initial work

#### Testing assumptions on ANOVA

```{r plotting fancily}
almost_sas <- function(aov.results){
  par(mfrow=c(2,2))
  plot(aov.results, which=1)
  plot(aov.results, which=2)
  aov_residuals <- residuals(aov.results)
  plot(density(aov_residuals))
  hist(aov_residuals)
}


almost_sas(wine_anova_results)
```




```
The residuals are normal, but variances are not equal. 
Since there are only two groups, should we look at doing an independent samples t-test with unequal variances?  Or is it better to run ANOVA/KW since we have so many samples?  I'm just curious why ANOVA was run instead of t-test
```

```{r}
y <- as.data.frame(wine_red)
p1 <- ggplot(y, aes(sample = total.sulfur.dioxide))
p1 <- p1 + 
        stat_qq() + 
        stat_qq_line() +
  labs(title="qq plot of sulfur dioxide in red wine")+
        theme_bw()

p1

y <- as.data.frame(wine_white)
p2 <- ggplot(y, aes(sample = total.sulfur.dioxide))
p2 <- p2 + 
        stat_qq() + 
        stat_qq_line() +
  labs(title="qq plot of sulfur dioxinde in white wine")+
        theme_bw()

p2
```

```
Both sets of data are not normal. Is this why we ran ANOVA?   
```
### Jovan's work

#### 2. Kruskal-Wallis Test on the Alcohol Content


```{r}
wine.Summary2 <-  wine %>% group_by(type) %>% summarise(alcohol_median = median(alcohol), alcohol_IQR = IQR(alcohol))
wine.Summary2
```

```{r}
wine_KW_results <- kruskal.test(alcohol ~ type, data = wine)
wine_KW_results
```

**Hypotheses**

$H_0: \ \mbox{M}_1 = \mbox{M}_2$ <br> 
$H_1: \ \mbox{M}_1 \ne \mbox{M}_2$ 

**Test Statistic**

$H = `r round(wine_KW_results$statistic,digits=2)`$.

***p*-value**

$`r p.value.string(wine_KW_results$p.value)`$.

**Rejection Region**

Reject if $p < \alpha$, where $\alpha=0.05$.

**Conclusion and Interpretation**

Fail to reject $H_0$. There is not sufficient evidence to suggest that there is a diferent in the alcohol contain of the two type of wines.

### Adrienne's work

#### Assumptions for KW

```{r}
alcohol_anova_results <- aov(alcohol ~ type, data = wine)
alcohol_anova_results

plot(alcohol_anova_results)

```
```{r}
almost_sas(alcohol_anova_results)
```

looking at theses graphs, we can see that while a lot of our residuals are on the 45 degree line, we have a lot that aren't.  The histogram of our residuals isn't quite bell shaped. It is skewed.  Therefore, we say that normality is broken and Kruskal-Walis was necessary.

```
Since we have only two groups, should we run an indpendent samples t-test? Or the nonparametric version?  Or is it better to run the anova and KW since we have so many samples?
```
### Jovan's work

#### 3. Test of variability pH


```{r}
pHvar <- var.test(wine_red$pH, wine_white$pH, alternative = "t", conf.level = 0.95, ratio = 1)
pHvar 
```

**Hypotheses**

* $H_0: \sigma_r = \sigma_w$
* $H_1: \sigma_r \ne \sigma_w$

**Test Statistic**

* $F_0 = `r round(pHvar$statistic, digits = 2)`$.
* $`r p.value.string(pHvar$p.value)`$.

**Rejection Region**

Reject $H_0$ if $p < \alpha$, where $\alpha=0.05$. $`r p.value.string(pHvar$p.value)` \nless 0.05.$

**Conclusion and Interpretation**

Fail to reject $H_0$. There is insufficient evidence to suggest that there is a difference in the variability of pH between red and white wines.


#### 4. Test of contain of Chlorides

Formally test to determine if the mean of the chlorides contain on red wine is greater than that of white wine. Use the appropriate version of the t-test with the critical at the α = 0.05 level.

```{r}
wine.Summary <-  wine %>% group_by(type) %>% summarise(mean_alcohol = mean(chlorides), stdev_alcohol = sd(chlorides))
wine.Summary
```

```{r}
chlorides_t <- t.test(wine_red$chlorides, wine_white$chlorides, paired = FALSE, alternative = "greater")
chlorides_t
chlorides_p <- p.value.string(chlorides_t$p.value)
```

**Hypotheses**

$H_0: \ \mu_R \le \mu_W$ <br> 
$H_1: \ \mu_R > \mu_w$

**Test Statistic**

$t_0 = `r round(chlorides_t$statistic, digits=3)`$.

***p*-value**

$`r chlorides_p`$.

**Rejection Region**

Reject if $p < \alpha$, where $\alpha=0.01$.

**Conclusion and Interpretation**

Reject $H_0$. There is sufficient evidence to suggest that the contain chlorides in the red wine are greather than the white wine.


#### 5. Test of variability of residual sugar 

```{r}
rsvar <- var.test(wine_red$residual.sugar, wine_white$residual.sugar, alternative = "t", conf.level = 0.95, ratio = 1)
rsvar
```

**Hypotheses**

* $H_0: \sigma_r = \sigma_w$
* $H_1: \sigma_r \ne \sigma_w$

**Test Statistic**

* $F_0 = `r round(rsvar$statistic, digits = 2)`$.
* $`r p.value.string(rsvar$p.value)`$.

**Rejection Region**

Reject $H_0$ if $p < \alpha$, where $\alpha=0.05$. $`r p.value.string(rsvar$p.value)` < 0.05.$

**Conclusion and Interpretation**

Reject $H_0$. There is sufficient evidence to suggest that there is a difference in the variability of residual sugar contain between red and white wines.


#### 6. Test of contain of Citrid Acid

Formally test to determine if the mean of the citrid acid contain on red wine is less than that of white wine. Use the appropriate version of the t-test with the critical at the α = 0.05 level.

```{r}
wine.Summary3 <-  wine %>% group_by(type) %>% summarise(mean_alcohol = mean(citric.acid), stdev_alcohol = sd(citric.acid))
wine.Summary3
```

```{r}
citricacid_t <- t.test(wine_red$citric.acid, wine_white$citric.acid, paired = FALSE, alternative = "less")
citricacid_t
citricacid_p <- p.value.string(chlorides_t$p.value)
```

**Hypotheses**

$H_0: \ \mu_R \ge \mu_W$ <br> 
$H_1: \ \mu_R < \mu_w$

**Test Statistic**

$t_0 = `r round(citricacid_t$statistic, digits=3)`$.

***p*-value**

$`r citricacid_p`$.

**Rejection Region**

Reject if $p < \alpha$, where $\alpha=0.01$.

**Conclusion and Interpretation**

Reject $H_0$. There is sufficient evidence to suggest that the contain of citric acid in the red wine is less than the white wine.


#### 7. Proportion Test on Density??? 

```
What are density units for wine?  What does this mean in the real world? What does density greater than one mean?  Why did we run this test three times with numbers so close to each other?  (not critisising... just trying to figure out what's going on)---this would be a good intro to this section
```

```{r}
wine_density1 <- z.test(wine$density, alternative = "greater", mu = 1, sigma.x = sd(wine$density))
wine_density1
```

**Hypotheses**

$H_0: \ \mu \le 1$ <br>
$H_1: \ \mu > 1$

**Test Statistic**

$z = `r wine_density1$statistic`$.

***p*-value**

$p = `r wine_density1$p.value`$.

**Rejection Region**

Reject if $p < \alpha$, where $\alpha=0.05$.

**Conclusion and Interpretation**

Fail to reject $H_0$. There is not sufficient evidence to suggest that the mean density wine is greater than 1.

```{r}
wine_density2 <- z.test(wine$density, alternative = "less", mu = 0.98, sigma.x = sd(wine$density))
wine_density2
```

**Hypotheses**

$H_0: \ \mu \ge 0.98$ <br>
$H_1: \ \mu < 0.98$

```
Changed the alternative hypothesis to match test run
```

**Test Statistic**

$z = `r wine_density2$statistic`$.

***p*-value**

$p = `r wine_density2$p.value`$.

**Rejection Region**

Reject if $p < \alpha$, where $\alpha=0.05$.

**Conclusion and Interpretation**

Fail to reject $H_0$. There is not sufficient evidence to suggest that the mean density wine is less than 0.98.

```{r}
wine_density3 <- z.test(wine$density, alternative = "less", mu = 0.99, sigma.x = sd(wine$density))
wine_density3
```

**Hypotheses**

$H_0: \ \mu \ge 0.99$ <br>
$H_1: \ \mu < 0.99$

```
Changed alternate hypothesis to match test run
```

**Test Statistic**

$z = `r wine_density3$statistic`$.

***p*-value**

$p = `r wine_density3$p.value`$.

**Rejection Region**

Reject if $p < \alpha$, where $\alpha=0.05$.

**Conclusion and Interpretation**

Fail to reject $H_0$. There is not sufficient evidence to suggest that the mean density wine is less than 0.99.


#### 8. Regression model

```
What is the regression actually modeling? for every one unit increase in all these predictors, what is our outcome?
Why don't we break it down and run just a few of these and have something that might be related?  like see if sulfur dioxide predicts pH or if alcohol content predicts sulfur 
```

Then in the idividual predictors, we need to test what we decide on one by one as predictors of our outcome.  

```{r}
one_model <- lm(quality ~ fixed.acidity + volatile.acidity + citric.acid + residual.sugar + chlorides + total.sulfur.dioxide + density + pH + sulphates + alcohol, data=wine)
one_coef <- coefficients(one_model)
one_anova <- anova(one_model)
one_summary <- summary(one_model)
one_t <- as_tibble(one_summary[[4]])
one_ci <- as_tibble(confint(one_model, level=0.95))
```

The resulting regression model is 
\[ \hat{y} = `r one_coef[[1]]` + `r one_coef[[2]]`x_1 + `r one_coef[[3]]`x_2 + `r one_coef[[4]]`x_3 + `r one_coef[[5]]`x_4 + `r one_coef[[6]]`x_5 + `r one_coef[[7]]`x_6 + `r one_coef[[8]]`x_7 + `r one_coef[[9]]`x_8 + `r one_coef[[10]]`x_9\]

#### ANOVA table and testing for significance of regression line

The corresponding ANOVA table is as follows

```{r}
one_anova
```

**Hypotheses**

$H_0: \ \beta_1=\beta_2=\beta_3=\beta_4=\beta_5=\beta_6=\beta_7=\beta_8=\beta_9 = 0$ <br>
$H_1: $ at least one is not zero

```
changed hypothesis to match test run.  The test is an F test... a multivariable model with nine betas. 
```


**Test Statistic**

$F_0 = `r round(one_anova$"F value"[1], digits=2)`$.

***p*-value**

$`r p.value.string(one_anova$"Pr(>F)"[3])`$.

**Rejection Region**

Reject if $p < \alpha$, where $\alpha=0.05$.

**Conclusion and Interpretation**

Comparing every p value of the previous table with the alpha, can be check that with at lest one of then is greather than alpha, by this reason fail to reject $H_0$. There is not sufficient evidence to suggest that the regression line is significant.

#### Testing for significance of individual predictor

##### Test for fixed acidity

**Hypotheses**

$H_0: \ \beta_1 = 0$ <br>
$H_1: \ \beta_1 \ne 0$

**Test Statistic**

$t_0 = `r round(one_t$"t value"[2], digits=2)`$.

***p*-value**

$`r p.value.string(one_t$"Pr(>|t|)"[4])`$.

**Rejection Region**

Reject if $p < \alpha$, where $\alpha=0.05$.

**Conclusion and Interpretation**

Comparing every p value of the previous table with the alpha, can be check that with at lest one of then is greather than alpha, by this reason fail to reject $H_0$. There is not sufficient evidence to suggest that fixed acidity is a significant predictor of 

-->
